# Results

## Plot 1
```{r}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)

ev_data <- read.csv('/Users/hujiasheng/Electric_Vehicle_Population_Data.csv')

ev_count_by_year <- ev_data %>% 
  group_by(Model.Year) %>%
  summarise(Count = n(), .groups= 'drop')

# Plotting the number of EVs registered each year
ggplot(ev_count_by_year, aes(x = Model.Year, y = Count)) +
  geom_line(na.rm = TRUE) +
  geom_point(na.rm = TRUE) +
  scale_x_continuous(limits = c(NA, 2023)) +
  theme_minimal() +
  ggtitle("Number of EVs Registered Each Year") +
  xlab("Year") +
  ylab("Number of EV Registrations")
```
**Number of EVs Registered Each Year (R-style Plot):** This plot shows the count of electric vehicles registered each year. The trend line and points provide a clear visualization of the growth in EV adoption over the years.

## Plot 2
```{r}
# Calculate year-over-year growth rate
ev_count_by_year <- ev_count_by_year %>%
  mutate(Growth = (Count / lag(Count) - 1) * 100)

# Plotting the year-over-year growth rate of EV registrations
ggplot(ev_count_by_year, aes(x = Model.Year, y = Growth)) +
  geom_line(color = "black", na.rm = TRUE) +
  geom_point(color = "black", na.rm = TRUE) +
  scale_x_continuous(limits = c(NA, 2023)) +
  theme_minimal() +
  ggtitle("Year-over-Year Growth Rate in EV Registrations") +
  xlab("Year") +
  ylab("Growth Rate (%)")
```
**Year-over-Year Growth Rate in EV Registrations (R-style Plot):** This plot represents the year-over-year growth rate in EV registrations, calculated as a percentage. It highlights the annual percentage change, indicating periods of significant growth or decline in EV adoption.
